---
title: "Analysis of Serial Measures"
author: "Holger Burchert"
date: "16 5 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This code is an implementation of the article "Analysis of Serial Measures in 
Medical Research" by Matthews et al. 1990: 

Matthews, J., Altman, D. G., Campbell, M. J., & Royston, P. (1990). Analysis of
serial measurements in medical research. British Medical Journal, 300(6719), 230-235.

Two groups (Group A and Group B) conducted an exercise test. The objective is 
to test if the heart rate of one group declines faster. The code conducts 
a linear regression on the heart rate vs time data for each participant. 
The slopes are then used to conduct a student's t-test to see if the 
slopes of the two groups differ significantly.

```{r} 
data <- read.csv(
    "C:/Users/neo/Documents/Privat/R/Summary Code/summary2.csv",
    header = TRUE,
    sep = ";"
  )
```

First some storages are built which will later be filled with outcomes from the
regressions. 

```{r}
# Group A suffix = .A 
 models.A   <- list() # storage for regression models  
 slopes.A   <- c()    # stroage for the slopes 
r.sqrds.A   <- c()    # storage for r squareds

# Group B suffix = .B
 models.B   <- list() # storage for regression models  
 slopes.B   <- c()    # stroage for the slopes 
r.sqrds.B   <- c()    # storage for r squareds
```

Now, a loop will be created. It conducts a regression von HR vs time and 
repeats this for each participant. During this loop, each regression model 
is stored in "models.A", while each slope is stored in "slopes.A" and the 
r-squared is stored in "r.sqrds.A". 

```{r}
# Looping regressions of HR vs time for Group A
for (i in names(data)[-1][1:9]) { # [-1] = omits time, [1:9] = desired columns         
   models.A[[i]] <-         lm(get(i) ~ zeit, data)                 # store model
   slopes.A[[i]] <-    coef(lm(get(i) ~ zeit, data))["zeit"]        # store slope
  r.sqrds.A[[i]] <- summary(lm(get(i) ~ zeit, data))[["r.squared"]] # store r^2
}
```

Since having the r-squareds is not enough (Anscombe's Quartett) all the
regressions for each participant are plotted to evaluate whether the linear
model is a good fit to the data. This is, again, done by looping the 
regressions. 

```{r}
par(mfrow = c(2, 3)) # merges the graphs to be looped in one graph
for (i in names(data)[-1][1:9]) { # [-1] = omits time, [1:9] = Group A 
  plot(get(i) ~ zeit,
         ylab = "HF (bpm)",
         main = i, 
          col = "steelblue",
          data) 
  abline(models.A[[i]])  # add linear fit from models
}
```

New the regressions are looped for group B

```{r}
# Looping regressions of HR vs time for Group B
for (i in names(data)[-1][10:15]) { # [-1] = omits time, [10:15] = Group B                   
   models.B[[i]] <-         lm(get(i) ~ zeit, data)                 # store model
   slopes.B[[i]] <-    coef(lm(get(i) ~ zeit, data))["zeit"]        # store slope
  r.sqrds.B[[i]] <- summary(lm(get(i) ~ zeit, data))[["r.squared"]] # store r^2
}
```

Now the regressions are plotted 

```{r}
# Looping regression plots for Group B
par(mfrow = c(2, 3)) # merges the looped graphs in one graph
for (i in names(data)[-1][10:15]) { # try indexing here as in other loops
  plot(get(i) ~ zeit,
         ylab = "HF (bpm)",
         main = i, 
         col  = "red",
         data)# Plotting
  abline(models.B[[i]])  # add linear fit from models
}
```

Creating histograms for the r-squareds to show how frequent certain values 
occured. 

```{r}
# Histogram r squared's Group A
par(mfrow = c(1,1))
hist(
  r.sqrds.A,
  freq   = TRUE,
  col    = "steelblue",
  xlab   = "r.squared",
  breaks = 25,
  main   = "Histogram Group A"
)

# Histogram r squared's Group B
par(mfrow = c(1,1))
hist(
  r.sqrds.B,
  freq   = TRUE,
  col    = "red",
  xlab   = "r.squared",
  breaks = 25,
  main   = "Histogram Group B"
)
```

## Conduct Statistical Tests

Assessing whether the slopes are Gaussian distributed. 

```{r}
# Assessing normal distribution Group A
par(mfrow = c(1,1)) # resets the "merge graphs" function
qqnorm(slopes.A,
       pch   = 1,
       frame = FALSE,
       main  = "Slopes.A")
qqline(slopes.A,
       col   = "steelblue",
       lwd   = 2)

# Assessing normal distribution Group B
par(mfrow = c(1,1)) # resets the "merge graphs" function
qqnorm(slopes.B,
       pch   = 1,
       frame = FALSE,
       main  = "Slopes.B")
qqline(slopes.B,
       col   = "red",
       lwd   = 2)
```

```{r}
# Assessing if variances are the same
# F test for equal variances 
var.test(slopes.A, slopes.B)
```

```{r}
# Conducting t-test
t.result <- t.test(
  slopes.A,
  slopes.B,
  paired      = FALSE,
  var.equal   = TRUE, # accroding to var.test result
  conf.level  = 0.95,
  alternative = "two.sided"
)

print(t.result)
```

Plotting the results

```{r}
boxplot(slopes.A, 
        slopes.B, 
        col = "steelblue",
        ylim = c(0.0,-0.2),
        main = "Results: Red lines indicate \n 95%CI diff from t-test"
)
abline(h = mean(slopes.B) + t.result$conf.int[1], col = "red") # 95%CI low. lim
abline(h = mean(slopes.B) + t.result$conf.int[2], col = "red") # 95%CI upp. lim
```